Heterogeneous warehouse fleets typically include robots with different roles, such as carrier vehicles and picker agents, that must coordinate task choice, role synchronization, and recharging while sharing shelves, corridors, and charging stations.
This creates a discrete multi-agent decision problem in which throughput depends not only on motion planning, but also on when agents should commit to tasks, wait for support, or allocate time to battery recovery.

This thesis evaluates whether off-the-shelf \glspl{llm} can support high-level decision-making for heterogeneous multi-robot collaboration in the Battery-TA-RWARE task-assignment warehouse environment.
The control loop summarizes the current environment state, queries an \gls{llm} for a grounded macro-action, and validates the proposed action against environment-defined feasibility constraints before execution.
The experiments compare model size, planning architecture, and prompt representation, using delivery throughput, \gls{llm} call efficiency, and observed failure behaviour as the main evaluation criteria.

The completed experiments show that, in the earlier shelf-delivery controller, shared-context planning performed better than centralized planning and natural-language prompting performed better than \gls{json} prompting.
In the later call-aware shared-context controller, a mid-sized local model (\nolinkurl{llama3.2:3b}) gave the strongest practical trade-off among the completed natural-language runs, while the completed \gls{json} runs remained substantially weaker than their natural-language counterparts.
Across the evaluated settings, productive behaviour depended on rule-based safeguards such as action validation and candidate shaping, indicating that \glspl{llm} were most useful as constrained high-level decision components rather than as unconstrained controllers.
