\section{Appendix}
\label{app:appendix}

\subsection{Reproducibility: Running the Controllers}
\label{app:repro}
The experiments in this thesis are executed from the Battery-TA-RWARE repository in this workspace.
Example commands:

\begin{lstlisting}[caption=Run the basic per-agent LLM controller for one episode., label=listing:run_llm]
python Battery-TA-RWARE/scripts/run_llm.py \
  --env_id tarware-tiny-1agvs-1pickers-partialobs-chg-v1 \
  --seed 0 \
  --num_episodes 1 \
  --model llama3.2:3b
\end{lstlisting}

\begin{lstlisting}[caption=Run the centralized shelf-delivery experiments., label=listing:run_centralized]
python Battery-TA-RWARE/scripts/run_centralized_llm_experiments.py \
  --prompt_format language \
  --seed 0
\end{lstlisting}

\begin{lstlisting}[caption=Run the shared-context shelf-delivery experiments., label=listing:run_shared_context]
python Battery-TA-RWARE/scripts/run_shared_context_llm_expt.py \
  --prompt_format language \
  --seed 0
\end{lstlisting}

\begin{lstlisting}[caption=Run the Objective~2 shared-context call-aware experiments., label=listing:run_obj2]
python Battery-TA-RWARE/scripts/run_obj2_shared_context_llm.py \
  --prompt_format language \
  --max_steps 0 \
  --max_busy_steps_for_replan 10 \
  --max_inactivity_steps 500 \
  --disable_support_needed_soon \
  --find_path_agent_aware_always \
  --seed 0
\end{lstlisting}

\begin{lstlisting}[caption=Run the heuristic policy for one episode., label=listing:run_heuristic]
python Battery-TA-RWARE/scripts/run_heuristic.py \
  --env_id tarware-tiny-1agvs-1pickers-partialobs-chg-v1 \
  --seed 0 \
  --num_episodes 1
\end{lstlisting}

The main result folders discussed in the thesis are:
\begin{itemize}
    \item \nolinkurl{results/shelf/central_llm_experiments}: Objective~1, centralized planning.
    \item \nolinkurl{results/shelf/shared_context_llm_experiments}: Objective~1, shared-context per-agent planning.
    \item \nolinkurl{results/calls/shared_context}: Objective~2, shared-context per-agent planning with call-efficiency objective.
    \item \nolinkurl{results/battery}: planned battery-oriented experiments (not yet populated with completed aggregate sessions at the time of writing).
\end{itemize}

\subsection{Prompt Excerpt (Illustrative)}
\label{app:prompt_excerpt}
The \gls{llm} is prompted with an explicit mapping from relevant state elements to discrete action identifiers.
The exact prompt depends on the experiment family: centralized prompts describe all agents at once, while shared-context prompts describe one current agent together with selected global context.
The full prompts are generated programmatically; the excerpt below illustrates the style (line breaks and ordering may vary):

\begin{lstlisting}[caption=Illustrative prompt excerpt., label=listing:prompt_excerpt]
Current global state:
- Requested shelves:
  - Requested shelf 24 is at position (11,5). To go there, choose action id 50.
  - Requested shelf 39 is at position (6,8). To go there, choose action id 29.
...
You are agent_0 (type=AGV).
Choose exactly one action_id from the candidate list.
Output format: action_id=<integer>
\end{lstlisting}

\subsection{Metric Definitions}
\label{app:metrics}
\begin{itemize}
    \item \textbf{Total shelf deliveries:} number of requested shelves delivered to goal locations during an episode.
    \item \textbf{Overall pick rate:} deliveries/hour computed as
    \[
        \text{pick\_rate} = \frac{\text{deliveries} \cdot 3600}{5 \cdot \text{episode\_steps}},
    \]
    where 5 steps/second is the environment's time scale used by the logging script.
    \item \textbf{Clashes and stuck events:} environment-defined counts related to collision resolution and deadlock-like behaviour.
    \item \textbf{Global episode return:} sum of environment rewards across all agents and time steps.
\end{itemize}
