This thesis explores \gls{llm}-driven decision-making for heterogeneous multi-robot collaboration in a task-assignment warehouse setting with battery constraints.
The work implemented an end-to-end control loop that (i) grounds decisions to discrete environment action identifiers, (ii) queries an \gls{llm} using centralized or per-agent shared-context prompts, and (iii) enforces feasibility through action-mask validation.

The results should be interpreted relative to a 20-delivery per-run ceiling in the current simulation setup. Against that ceiling, Objective~1 established the initial shelf-delivery comparison between architectures and prompt forms. In that earlier controller family, shared-context planning was more promising than centralized planning, and natural-language prompting was more effective than \gls{json} prompting. These results were informative, but they remained far below the achievable ceiling and therefore served more as an indication of architectural direction than as evidence of strong final performance.

Objective~2 then changed the optimization target from pure shelf delivery to call-aware shelf delivery while also using a later, safer shared-context runner. The completed Objective~2 sessions show that the framework remained usable under the changed objective, especially in the more structured shared-context setup. Within that family, model size mattered, but not monotonically: larger models were not uniformly better, and the strongest practical trade-off in the natural-language session came from a mid-sized model, \texttt{llama3.2:3b}, rather than from the largest available model. The completed Objective~2 \gls{json} session produced non-zero throughput, but it remained substantially weaker than natural language in both throughput and call efficiency.

Objective~3, the battery-aware objective, remains open because completed aggregate result files were not yet available. This means the thesis only partially answers the broader question of whether the framework adapts smoothly as the objective changes. The evidence supports viability across the shift from Objective~1 to Objective~2, but not yet for a dedicated battery-centered objective.

The strongest overall lesson is not that the \gls{llm} replaced controller logic, but that it contributed best as a constrained macro-action selector inside a safeguarded system. Many poor runs ended in long non-productive tails, often far below the 20-delivery ceiling and usually without clear low-level collision deadlocks. This suggests that higher-level coordination did not become reliable simply by inserting an \gls{llm} into the loop. Action validation, candidate shaping, conflict handling, and other rule-based logic remained necessary to keep the system stable and productive.

The overall evidence in this thesis therefore supports a constrained view of \gls{llm} control. In the evaluated warehouse setting, \glspl{llm} were most useful inside a strongly structured decision loop, while correctness-critical behaviour remained better handled by explicit rule-based logic.

\subsection{Future Work}
\label{sec:futurework}
\input{text/futurework}
