In studies of human communication, language use has been framed as a form of joint action in which participants coordinate their individual contributions toward shared outcomes~\cite{clark1996usinglanguage}.
Accounts of language origins likewise emphasize the role of cooperative motives and shared intentionality in enabling coordination and collaboration~\cite{tomasello2008origins}.
Language has also been linked to social bonding and group-level coordination in evolutionary accounts of human sociality~\cite{dunbar1996grooming}.

Motivated by these perspectives, this thesis adopts the working hypothesis that language-capable models with general reasoning ability can serve as high-level decision-makers in multi-agent settings, thereby supporting collaboration and coordination without task-specific training for every new configuration.

To evaluate this hypothesis in a concrete and measurable way, the thesis uses a simplified but structured multi-agent testbed.
When autonomous agent fleets operate in a shared environment, high throughput typically requires more than path planning: agents must decide \emph{which} tasks to execute, \emph{when} to coordinate with other roles, and \emph{when} to allocate time to maintenance actions such as recharging.
These decisions are particularly challenging in heterogeneous teams where agents have different capabilities and constraints.

Research on \glspl{mrs} spans architectures, communication, and coordination strategies.
Early surveys emphasized fundamental axes such as group architecture, resource conflicts, and learning~\cite{cao1997cooperative}.
Later overviews categorized multi-robot research topics including communication, task allocation/control, and motion coordination~\cite{arai2002advances}.
In practical logistics domains, these themes manifest as repeated, interdependent choices: allocate requests to carriers, ensure loaders and carriers meet at shelves, resolve conflicts in shared corridors, and maintain energy safety.

Many approaches address such problems using optimization, hand-engineered heuristics, or \gls{marl}.
Optimization and heuristics can be strong and efficient, but they often require careful domain-specific design.
\gls{marl} can learn sophisticated joint behaviour, but training is expensive, sensitive to reward shaping and partial observability, and may generalize poorly when the environment parameters change.
Benchmarking studies highlight that algorithm comparisons are highly task- and metric-dependent, motivating careful protocol design and reporting~\cite{papoudakis2020benchmarking}.
Recent work in heterogeneous multi-robot settings also introduces bio-inspired \emph{symbiosis} framings and reward shaping strategies, often combined with centralized training and decentralized execution, to stabilize learning and improve cooperative outcomes~\cite{niu2025icpssymbiosis,niu2025icrassymbiosisrewards}.

Recent \glspl{llm} offer a complementary direction: instead of learning a policy from scratch for each environment, an \gls{llm} can be prompted with a structured description of the current state and asked to output a high-level decision.
In robotics, \glspl{llm} have been used for instruction following and planning by combining language reasoning with explicit grounding and feasibility checks; for example, by selecting from validated affordances rather than unconstrained free-form actions~\cite{ahn2022do}.
However, applying \glspl{llm} to multi-agent control introduces additional challenges: actions must remain valid under environment constraints, decisions must be coordinated across agents, and long-horizon objectives must be approximated under limited context and latency budgets.

This thesis studies \gls{llm}-driven decision-making in a heterogeneous multi-robot warehouse domain with explicit battery dynamics.
The experimental platform is Battery-TA-RWARE, an adaptation of the Task-Assignment Robotic Warehouse environment~\cite{krnjaic2022scalable}, featuring two agent types: carrier \glspl{agv} and picker agents that must meet at shelf locations for load/unload operations.
At each decision point, the current global state is summarized into a compact textual prompt, the \gls{llm} proposes an action id per agent, and a validation layer enforces action feasibility via action masks.

\subsection{Research Questions}
\label{sec:rq}
The thesis is guided by the following main research question:
\begin{quote}
    \textbf{Main research question:} How can \glspl{llm} be used in heterogeneous multi-robot systems to improve collaboration?
\end{quote}
To support a systematic investigation of the main research question, the study considers two sub-questions:
\begin{enumerate}
    \item \textbf{RQ1 (Model size):} How do \glspl{llm} of different sizes influence the quality and effectiveness of collaboration in heterogeneous multi-robot systems?
    \item \textbf{RQ2 (Integration architecture):} Which integration architecture, centralized or decentralized, enables the most effective use of \glspl{llm} for coordinating heterogeneous multi-robot teams?
\end{enumerate}

In this thesis, architectural choices include how environment state is translated into language, how proposed decisions are mapped back to executable actions, whether decision-making is performed per-agent or jointly, and how feasibility constraints are imposed (e.g., by restricting/validating candidate actions and using environment feedback for replanning).

\subsection{Contributions}
\label{sec:contrib}
The current thesis work contributes:
\begin{itemize}
    \item An end-to-end pipeline for running and evaluating \gls{llm}-driven heterogeneous multi-robot control in Battery-TA-RWARE, including state-to-language encoding, action-id grounding, output parsing, and feasibility validation.
    \item An experimental setup for comparing models of different sizes and for evaluating centralized versus decentralized \gls{llm} integration patterns.
    \item Preliminary results and an error analysis comparing different \gls{llm} planning architectures, prompt forms, and objective formulations.
\end{itemize}

\subsection{Report Outline}
Section~\ref{sec:purpose} states the problem formulation, objectives, and delimitations.
Section~\ref{sec:background} provides background on multi-robot coordination, warehouse task assignment, and \gls{llm}-based planning.
Section~\ref{sec:method} describes the environment and the proposed \gls{llm} control loop.
Section~\ref{sec:result_discussion} presents preliminary experiments and discusses observed failure modes.
Finally, Section~\ref{sec:conclusions} concludes and outlines future work.
