\subsection{Overview}
This chapter serves two roles.
It describes the implemented system, but it also justifies the empirical design used to answer the research questions.
The thesis is therefore structured as a comparative experimental study: the same warehouse environment, evaluation metrics, and scenario catalogue are reused while selected design variables are changed in order to study their effect on collaborative task performance.

This design is appropriate for the thesis because the research questions are comparative rather than purely descriptive.
RQ1 asks how model size affects performance, RQ2 asks how planning architecture affects performance, and RQ3 asks how prompt representation affects performance.
These questions require the report to make explicit what is fixed, what is varied, and how the experiments are structured across the objective families.

The thesis uses an empirical, experimental methodology:
\begin{enumerate}
    \item Implement an \gls{llm}-driven control loop for a simulated heterogeneous warehouse environment.
    \item Define the main comparison axes and evaluation metrics.
    \item Run controlled experiments (fixed environment ids and seeds) across multiple \gls{llm} models, planning architectures, and prompt variants.
    \item Analyze throughput, safety, and collaborative task outcomes as well as qualitative failure modes.
\end{enumerate}

\paragraph{What is fixed and what is varied.}
Across the main experiment families, the environment, action space, scenario catalogue, evaluation metrics, grounding/validation scheme, and local-inference setup are held as stable as practical.
The principal variables are model size, planning architecture, prompt representation, and objective family.
The report does not use an external heuristic baseline in the current comparison set, so the reported comparisons are made within the language-model-based control framework.

\paragraph{Interpretation and controls.}
The experiments are controlled in the sense that scenario ids, seeds, and runner configurations are recorded and reused within each family.
The surrounding validation layer filters infeasible outputs before execution, and the same comparison axes are used across the objective families.

\subsection{Environment: Battery-TA-RWARE}
Battery-TA-RWARE is a discrete-time, grid-based warehouse simulation derived from \gls{tarware}~\cite{krnjaic2022scalable}.
At any time, a set of shelves is requested. Carrier \glspl{agv} must transport requested shelves to goal locations and then return them to storage.
Picker agents are required to meet \glspl{agv} at shelf locations for load and unload interactions, creating explicit heterogeneous coordination constraints.
Agents additionally have a battery state that decreases with movement and interactions; charging stations are available at fixed locations.

The environment exposes a discrete action space where each action corresponds to selecting a target location:
\begin{itemize}
    \item \textbf{Shelf actions:} choose a shelf location (e.g., to pick up or return a shelf).
    \item \textbf{Goal actions:} choose a delivery station (primarily relevant when an \gls{agv} carries a requested shelf).
    \item \textbf{Charging actions:} choose a charging station to recover battery.
\end{itemize}
Low-level motion toward the selected target is handled internally by the environment (A* path planning and collision resolution), while the controller is evaluated on the quality of these high-level decisions.

\subsection{Language-Model Control Loop}
\label{sec:llm_loop}
The \gls{llm} policy is implemented as a text-in / discrete-action-out controller:
\begin{enumerate}
    \item \textbf{State summarization:} extract salient features of the global state (requested shelves and their action ids, per-agent position/target/battery/carrying status, charging occupancy).
    \item \textbf{Candidate action set:} compute a valid-action mask from the environment and derive the prompt-visible candidate set from those feasible actions. These candidates are presented as compact role-relevant action rows with action ids and guidance fields such as distance, tags, and charging-related instructions.
    \item \textbf{Prompting:} generate a per-agent prompt describing the agent role and the current state, including explicit mappings from requested objects to action ids.
    \item \textbf{Querying:} send the prompt to a local inference server (Ollama) and obtain a text response.
    \item \textbf{Parsing:} parse an integer action id (and optionally an action hold duration) from the response.
    \item \textbf{Validation:} validate the parsed action id against the current valid-action mask. If the proposed action is missing or invalid, the controller discards it and falls back to an idle action for that step; in the action-persistence setting, an invalid persisted action clears the hold and forces a new query.
    \item \textbf{Execution:} apply the validated joint action in the environment for one step (or hold the action for multiple steps, depending on configuration).
\end{enumerate}

\begin{figure}[H]
    \centering
    \begin{tikzpicture}[
        node distance=5mm,
        box/.style={draw, rounded corners, align=center, minimum width=0.7\textwidth, minimum height=8mm},
        arrow/.style={-Latex, thick}
    ]
        \node[box] (state) {State summary from the current environment step};
        \node[box, below=of state] (candidates) {Candidate shaping and feasible-action mask};
        \node[box, below=of candidates] (prompt) {Prompt construction for joint or per-agent planning};
        \node[box, below=of prompt] (query) {Language-model query};
        \node[box, below=of query] (parse) {Parse action identifier and optional hold duration};
        \node[box, below=of parse] (validate) {Validate against current constraints};
        \node[box, below=of validate] (execute) {Execute one environment step or continue a persisted action};
        \node[box, below=of execute] (feedback) {Observe step consequence and decide whether replanning is needed};
        \draw[arrow] (state) -- (candidates);
        \draw[arrow] (candidates) -- (prompt);
        \draw[arrow] (prompt) -- (query);
        \draw[arrow] (query) -- (parse);
        \draw[arrow] (parse) -- (validate);
        \draw[arrow] (validate) -- (execute);
        \draw[arrow] (execute) -- (feedback);
        \draw[arrow] (feedback.east) -- ++(2.2,0) |- (state.east);
    \end{tikzpicture}
    \caption{Schematic of the language-model control loop used in the experiments.}
    \label{fig:lm_control_loop}
\end{figure}

The controller is best described as a discrete-time, step-synchronous supervisory loop with conditional replanning.
The environment advances one simulation step at a time, and at each step the controller reevaluates agent state and action validity.
However, this does not imply that every agent is queried at every step.
Instead, the experiments use two related control schemes.

\begin{itemize}
    \item \textbf{Step-synchronous replanning with environment-owned continuation:} agents that are currently available for planning may receive a new high-level action, while busy agents are not replanned and continue their current task until the environment releases them.
    \item \textbf{Step-synchronous replanning with action persistence and conditional override:} accepted actions may remain active for multiple future steps, so agents are not re-queried while the persisted action remains valid. Replanning is triggered when a persisted action expires, becomes invalid, or when the controller decides intervention is needed, including selected busy-agent cases.
\end{itemize}

This design sits between two extremes.
A fully scheduled scheme that queries every agent at every step would generate many unnecessary \gls{llm} calls, while a purely event-triggered scheme is less natural in this setting because useful triggers such as blocked progress, invalid persisted actions, expiring commitments, or changing support needs are only known after the system state is checked.
The adopted design therefore combines regular per-step observability with selective replanning.
This is also consistent with practical supervisory control in warehouse systems, where the fleet is typically checked at regular intervals, ongoing tasks continue by default, and new task-level decisions are issued only when the current situation justifies intervention.

Two design choices are central for feasibility:
\begin{itemize}
    \item \textbf{Grounding via action ids:} the \gls{llm} does not output coordinates or free-form text commands; instead it selects from discrete action ids that are already defined by the environment.
    \item \textbf{Hard validation:} action masks enforce constraints such as invalid goals for picker agents and busy-state restrictions. Invalid \gls{llm} outputs are never executed.
\end{itemize}

The reported behaviour is therefore the behaviour of the complete control framework: prompt design, model output, candidate shaping, validation logic, and replanning policy act together inside the same supervisory loop.

\subsection{Experiment Families and Objectives}
The implemented experiments are grouped into three objective families.
Across these families, the common control framework is the grounded language-model loop described above: step-synchronous execution, conditional replanning, explicit action validation, and action persistence when configured.
Across all three objective families, the same controlled variables are intended: planning architecture, prompt representation, model size, scenario catalogue, and evaluation metrics.
What changes between the families is the objective function.

\paragraph{Objective 1: maximize total shelf deliveries.}
The first family fixes the operational objective to raw shelf-delivery throughput.
This family provides the delivery-only baseline against which the other objective formulations are interpreted.

\paragraph{Objective 2: optimize \gls{llm} calls while maintaining delivery throughput.}
The second family changes the objective from delivery-only to a joint objective: increase shelf deliveries while reducing unnecessary \gls{llm} queries.
It primarily supports RQ1 and the call-efficiency analysis while keeping the same control framework and the same comparison axes as Objective~1.
In this family, the output interface includes a \nolinkurl{Steps} field so that the selected action can remain active for multiple steps when that serves the call-aware objective.

\paragraph{Objective 3: battery-aware objective.}
A third family is defined for explicitly battery-oriented evaluation.
Its role is to extend the same framework toward constraint-aware behaviour beyond throughput and call efficiency.

\subsection{Planning Architectures}
The experiments compare two distinct ways of integrating the \gls{llm} into the decision loop.

\paragraph{Centralized planning.}
In the centralized architecture, one \gls{llm} call plans all agents that currently need a decision. This design exposes the full global state and all relevant candidate actions in one prompt. Its intended advantage is explicit joint coordination. Its main risk is prompt overload: as the number of agents and candidate actions increases, the prompt becomes large and more difficult to parse reliably.

\paragraph{Shared-context per-agent planning.}
In the shared-context architecture, the \gls{llm} plans one agent at a time. Each prompt still includes global information, but the final decision concerns only the current agent. This reduces the output burden and simplifies parsing, at the cost of requiring coordination to emerge from repeated local decisions rather than from one explicit joint plan. In the thesis evaluation, this architecture is the main vehicle for comparing prompt forms and model sizes under a common decision structure.

\subsection{Prompt Structures}
The prompt comparison in this thesis varies along two axes:
\begin{itemize}
    \item \textbf{Planning scope:} centralized joint planning versus per-agent planning with shared context.
    \item \textbf{Output structure:} natural-language responses versus \gls{json} responses.
\end{itemize}

The prompts include:
\begin{itemize}
    \item explicit requested-shelf to action-id mappings,
    \item role-specific rules for \glspl{agv} and pickers,
    \item support relations such as waiting-for-picker context,
    \item battery-state abstractions,
    \item candidate action filtering or prioritization,
    \item and, in Objective~2, an explicit \nolinkurl{Steps} field as part of the call-aware output interface for reducing unnecessary \gls{llm} calls.
\end{itemize}

\subsection{Comparison Axes}
The main comparison axes are:
\begin{itemize}
    \item model size,
    \item planning architecture (centralized versus shared-context per-agent),
    \item prompt form (natural language versus \gls{json}),
    \item and objective formulation (shelf-delivery only, call-aware shelf delivery, and battery-aware control) under the same grounded control framework.
\end{itemize}

\subsection{Evaluation Metrics}
The main metric is delivery throughput, expressed as a pick rate (deliveries per hour) derived from the number of delivered shelves and the simulated episode duration.
Additional metrics are collected to capture coordination and safety:
\begin{itemize}
    \item total shelf deliveries,
    \item number of clashes and deadlocks/stuck events,
    \item distance travelled by each agent type,
    \item episode return (environment reward),
    \item episode length and inference speed/latency (when measured).
\end{itemize}
The primary purpose of these metrics is to evaluate whether the language-model-based controller adapts productively when the objective changes.
Throughput remains the central outcome measure, since each objective still depends on productive task completion.
The additional metrics are used as supporting evidence for how that adaptation occurs in practice, for example whether a configuration maintains delivery performance, reduces unnecessary model calls, or exhibits coordination failures such as clashes, deadlocks, zero-delivery runs, or long no-delivery tails.

\subsection{Scenario and Model Matrix}
The same scenario catalogue is reused across the experiment families:
\begin{itemize}
    \item \nolinkurl{tiny_balanced_1v1},
    \item \nolinkurl{small_balanced_2v2},
    \item \nolinkurl{medium_balanced_4v4},
    \item \nolinkurl{small_agv_heavy_4v2},
    \item \nolinkurl{medium_agv_heavy_6v2},
    \item \nolinkurl{small_picker_heavy_2v4}.
\end{itemize}
These scenarios were selected to probe different coordination pressures: simple balanced collaboration, increased concurrency, larger warehouses, \gls{agv}-heavy support bottlenecks, and picker-heavy underutilization.
More specifically, the balanced small scenarios provide lower-pressure comparison points for model and prompt behaviour, the larger balanced scenarios expose increased concurrency and scalability pressure, the \gls{agv}-heavy scenarios stress picker-support bottlenecks and waiting dependencies, and the picker-heavy scenarios stress allocation efficiency and underutilization.
The scenario set is used to test whether the same model or architecture remains productive when the source of coordination difficulty changes.

The model set spans smaller and larger local models:
\begin{itemize}
    \item \nolinkurl{llama3.2:1b},
    \item \nolinkurl{llama3.2:3b},
    \item \nolinkurl{mistral:7b},
    \item \nolinkurl{gemma3:12b},
    \item \nolinkurl{phi4},
    \item \nolinkurl{qwen2.5:14b}.
\end{itemize}

\subsection{Experimental Protocol}
The reported experiment sessions were run with fixed environment ids and a fixed seed.
This defines a controlled set of comparable runs for the reported objective families.

The progression of the experimental campaign was:
\begin{enumerate}
    \item Objective~1 shelf-delivery experiments,
    \item Objective~2 call-aware shelf-delivery experiments under the same control framework and comparison axes,
    \item Objective~3 battery-aware experiments.
\end{enumerate}

Section~\ref{sec:result_discussion} reports Objective~1 for the shelf-delivery comparisons, Objective~2 for the call-aware comparisons, and Objective~3 for the battery-aware objective.
In the current simulation setup, a run can achieve at most 20 deliveries, and the results chapter uses that ceiling to contextualize reported scores.

\subsection{Delimitations}
To keep the study focused and interpretable, the following delimitations apply:
\begin{itemize}
    \item The work is evaluated in a simulated warehouse environment rather than on physical robots.
    \item The controller is responsible only for high-level target selection; low-level navigation and collision handling are performed by the environment.
    \item The evaluated language models are prompted general-purpose models accessed through a local inference server and are not fine-tuned for this task.
    \item The reported experiments use fixed seeds and a limited number of completed sessions.
    \item Completed aggregate results are not yet available for the battery-aware objective family.
\end{itemize}

\subsection{Ethical and Sustainability Considerations}
Although the present work is simulation-based, it raises practical issues that remain relevant for real multi-robot deployments.
The primary ethical concern is safety and reliability: generated decisions may be infeasible, unstable, or unproductive if they are not constrained by system-side checks.
For that reason, the studied control loop does not execute free-form outputs directly. It uses grounded action identifiers, candidate shaping, explicit feasibility validation, and fallback behavior so that unsafe or clearly invalid actions are filtered before execution.
The logging of prompts, parsed outputs, executed actions, and observed step consequences also improves traceability, which is important when analyzing failures such as long no-delivery tails, stale waiting loops, or battery-related breakdowns.

From a sustainability perspective, the main trade-off is between model capability and computational cost.
Larger language models generally require more inference time and compute, while smaller models may require stronger prompt structure and validation to remain effective.
This thesis therefore treats call efficiency, unnecessary replanning, and coordination overhead as practically relevant concerns rather than only implementation details.
Even with safeguards, fixed-seed evaluation, limited scenario coverage, and the absence of physical deployment still bound the scope of the reported results.
